---
title: "Genomicper Notebook - 50k Use case"
author: " Aybuke Ozcelik, Irem Okur, Mario Antonioletti, Pau Navarro, Claudia Cabrera"
output: 
  html_document:
    toc: true
  md_document:
      variant: markdown_github
      toc: true
editor_options: 
  markdown: 
    wrap: 72
---

## Setup

```{r setup, echo=FALSE, message=FALSE, warning = FALSE}

library(genomicper)        # Package to be analysed
library(microbenchmark)    # Microbenchmark routines
library(profvis)           # Profile routines
library(lintr)             # Static R analyser
library(diffr)             # Source file comparison
library(here)              # Determine the root of where the files live
library(compiler)          # Library to compiler byte compile R functions
library(testthat)          # Testing routines.
library(ggplot2)           # Add plots
library(dplyr)             # Manipulate data frames
library(readr)             # tidyverse reading tools
library(pryr)              # Memory usage
library(vroom)
library(data.table)
# Make sure the working directory is set to where this file lives
# across code chunks.
dir <- paste0(here(), "/Workspaces/")
knitr::opts_knit$set(root.dir = dir)

# Remove all objects in the current environment
rm(list = ls())

```

## Baseline performance

### Load functions

Find out how long routines are taking to run. To start we need to load up the
new version of the files. If you change these you will have to load them up 
again. You should probably focus on the routines that are taking the longest 
time.

```{r read_alternative_files}

# Load all the alternative versions of the code (follow the *_v2.R pattern)
new_files <- list.files(path = "Src", pattern = "*_v2.R", full.names = TRUE)

# Check the files have been read in.
if(length(new_files) == 0){
  stop("Error: not read in any files.")
}

# Loop round the files and source them to load them to the environment
for (file in new_files){
  source(file = file)
}

```

## Profile

Profile an example piece of code:

```{r profile_50k}

profvis({

    #### SNP-level  ##########################################################
    # SNPs annotation and Pathways provided by user
    # all data stored at the WORKSPACE

    # Read & format GWAS pvalues
    all_data <- fread(file = "Data/AllData50k.tsv",data.table = FALSE)
    
    # Order data according to the genome
    genome_results <- genome_order_v2(all_data = all_data)

    # Results from genome_order
    ordered_alldata <- genome_results$ordered_alldata
    gs_locs <- genome_results$gs_locs

    # Create new environment to save variables (e.g. pathways, permutations):
    gper.env <- new.env()

    # Pathways can be downloaded using the function get_pathways()
    # Load example pathways into the new environment.
    data(RHSA164843, RHSA446343, RHSA8876384, RHSA8964572,
         RHSA109582, RHSA1474244, envir = gper.env)

    # Map SNPs to pathways
    paths_res <- read2_paths_v2(ordered_alldata = ordered_alldata,
                             gs_locs = gs_locs,
                             sets_from = "workspace",
                             sets_prefix = "RHSA",
                             level = "snp",
                             envir = gper.env)

    # Results from read2_paths:
    pers_ids <- paths_res$per_ors
    pathways <- paths_res$pathways

    # Perform permutations:
    # nper should be bigger!
    snps_permutation_v2(ordered_alldata = ordered_alldata,
                     pers_ids = pers_ids, 
                     ntraits = 7, 
                     nper = 10, 
                     saveto = "workspace",
                     threshold = 0.05, 
                     gs_locs = gs_locs, 
                     envir = gper.env)

    # Get results
    results <- get_results_v2(res_pattern = "Permus", level = "snp",
                           from = "workspace", threshold = 0.05,
                           envir = gper.env)

  
    })

```


## Comparsion of vroom vs fread - 50K use case

```{r reading_data_50K}

cols <- cols(
  name = col_character(),
  Trait = col_double()
)

res <- microbenchmark(
                      vroom(file = "Data/AllData50k.tsv"),
                      fread(file = "Data/AllData50k.tsv"),
                      times = 20
                    )

# Boxplot of the results
boxplot(res, xlab = "vroom vs fread", 
        names = c("vroom", "fread"),
        col = c("red", "yellow"))

# Using ggplot explicitly to plot the results
res %>% select(time, expr)         %>% 
        mutate(time = time / 10^9) %>% # Convert to seconds
        ggplot(aes(x = expr, y = time, fill = expr), alpha = 0.5) +
        geom_violin() +
        theme(legend.position = "none", 
              plot.title = element_text(hjust = 0.5)) +
        ylab("Time (s)") + xlab("Function version") +
        ggtitle("Comparsion of vroom vs fread for 30K entries") +
        scale_x_discrete(labels = c("vroom", "fread")) 

# Check memory usage, see:
# http://adv-r.had.co.nz/memory.html

gwas <- vroom(file = "Data/AllData50k.tsv")
tgwas <- fread(file = "Data/AllData50k.tsv")
object_size(gwas)   # On Irem's laptop 2,208 B, 4,144 B on Mac
object_size(tgwas)  # On Irem's laptop 2,399,304 B, same on Mac


```

### Genome_order changes

```{r prof_genome_order}

profvis({
  
    # Order data according to the genome using v2 makes it possible to make
    # changes to the code base to try an improve the performance.
    all_data <- fread(file = "Data/AllData50k.tsv",data.table = FALSE)
    genome_results <- genome_order_v2(all_data = all_data)

})


```

```{r microbenchmark_genomic_order}

# Run only 5 times if it takes a long time to run
gres <- microbenchmark(
  genome_order(all_data = all_data),
  genome_order_v2(all_data = all_data),
  unit = "s",
  times = 20
)

# print out the results
print(gres)

# Boxplot of the results
boxplot(gres, 
        xlab = "Function", 
        names = c("genome_order", "genome_order_v2"),
        col = c("red", "yellow"))
#
gres %>% select(time, expr)         %>% 
        mutate(time = time / 10^9) %>% # Convert to seconds
        ggplot(aes(x = expr, y = time, fill = expr), alpha = 0.5) +
        geom_violin() + scale_y_log10() +
        theme(legend.position = "none", 
              plot.title = element_text(hjust = 0.5)) +
        ylab("Time (s)") + xlab("Function version") +
        ggtitle("Comparsion of genome_order improvements for 50k entries") +
        scale_x_discrete(labels = c("genome_order", "genome_order_v2")) +
        theme_bw() + theme(legend.position = "none")

```
```{r mem_profile_genome_order}

library(profmem)

profmem({
    # Order data according to the genome using v2 makes it possible to make
    # changes to the code base to try an improve the performance.
    genome_results <- genome_order_v2(all_data = all_data)
})
```

