---
title: "Genomicper - 30M usecase"
author: "Mario Antonioletti, Pau Navarro, Claudia Cabrera"
output: 
  html_document:
    toc: true
  md_document:
      variant: markdown_github
      toc: true
editor_options: 
  markdown: 
    wrap: 72
---

## Setup

```{r setup, echo=FALSE, message=FALSE, warning = FALSE}

library(genomicper)        # Package to be analysed
library(microbenchmark)    # Microbenchmark routines
library(profvis)           # Profile routines
library(lintr)             # Static R analyser
library(diffr)             # Source file comparison
library(here)              # Determine the root of where the files live
library(compiler)          # Library to compiler byte compile R functions
library(testthat)          # Testing routines.
library(ggplot2)           # Add plots
library(dplyr, warn.conflicts = FALSE) # Manipulate data frames
library(readr)             # tidyverse reading tools
library(pryr)              # Memory usage
library(data.table)        # For the fread function
library(vroom)             # To read files
library(testthat)          # Testing routines

# Make sure the working directory is set to where this file lives
# across code chunks.
dir <- paste0(here(), "/Workspaces/")
knitr::opts_knit$set(root.dir = dir)

# Remove all objects in the current environment
rm(list = ls())

```

## Baseline performance

### Load functions

Find out how long routines are taking to run. To start we need to load up the
new version of the files. If you change these you will have to load them up 
again. You should probably focus on the routines that are taking the longest 
time.

```{r read_alternative_files, warning=TRUE}

# Load all the alternative versions of the code (follow the *_v2.R pattern)
new_files <- list.files(path = "Src", pattern = "*_v2.R", full.names = TRUE)

# Check the files have been read in.
if(length(new_files) == 0){
  stop("Error: not read in any files.")
}

# Loop round the files and source them to load them to the environment
for (file in new_files){
  source(file = file)
}

```

## Pofile

```{r benchmark_reading_data}

# Column types for read_table
cols <- cols(
  name = col_character(),
  Chromosome = col_double(),
  Location = col_double(),
  GENE_ID = col_double(),
  Symbol = col_character(),
  Orientation = col_character()
)

# 30M file to read in
infile <- "Data/SNP_Table_Annotation_ALL.txt.gz"

# Perform the benchmark
res <- microbenchmark(
                      read.table(file = infile,
                                 header = TRUE, 
                                 stringsAsFactors = FALSE),
                      read_table(file = infile,
                                 col_types = cols),
                      fread(file = infile),
                      times = 5
                    )

# Print out the stats
print(res)

# Unit: seconds
#                                                                expr      min       lq     mean   median        uq
#  read.table(file = infile, header = TRUE, stringsAsFactors = FALSE) 93.55874 95.22300 99.07242 97.39790 102.82382
#                         read_table(file = infile, col_types = cols) 56.47857 57.23174 63.71495 65.56512  67.89029
#                                                fread(file = infile) 30.55904 33.09605 34.83237 35.52654  36.67081
#        max neval
#  106.35863     5
#   71.40905     5
#   38.30937     5

# Boxplot of the results
boxplot(res, xlab = "read.table vs read_table vs fread", 
        names = c("read.table", "read_table", "fread"),
        col = c("red", "yellow"))

# Using ggplot explicitly to plot the results
res %>% select(time, expr)         %>% 
        mutate(time = time / 10^9) %>% # Convert to seconds
        ggplot(aes(x = expr, y = time, fill = expr), alpha = 0.5) +
        geom_violin() +
        theme(legend.position = "none", 
              plot.title = element_text(hjust = 0.5)) +
        ylab("Time (s)") + xlab("Function version") +
        ggtitle("Comparsion of read.table vs read_table for 30M entries") +
        scale_x_discrete(labels = c("read.table", "read_table", "fread")) 

# Check memory usage, see:
# http://adv-r.had.co.nz/memory.html

gwas <- read.table(file = infile,
                   header = TRUE, 
                   stringsAsFactors = FALSE)

tgwas <- read_table(file = infile,
                           col_types = cols)

fgwas <- fread(file = infile)

vgwas <- vroom(file = infile, col_types = cols)

object_size(gwas)   # On MA's Mac 3,287,884,016 B ~ 3GB - data table
object_size(tgwas)  # On MA's Mac 3,726,541,312 B ~ 3.47Gb -tibble
object_size(fgwas)  # On MA's Mac 3,287,884,328 B
object_size(vgwas)  # On MA's Mac 5,736 B (?!)


# Test load time between vroom and fread
res3 <- microbenchmark(
                      vroom(file = infile,
                                 col_types = cols),
                      fread(file = infile),
                      times = 10
                    )

# Print out the stats
print(res3)

# Boxplot of the results
boxplot(res3, xlab = "vroom vs fread", 
        names = c("vroom", "fread"),
        col = c("red", "yellow"))

# Using ggplot explicitly to plot the results
res3 %>% select(time, expr)         %>% 
        mutate(time = time / 10^9) %>% # Convert to seconds
        ggplot(aes(x = expr, y = time, fill = expr), alpha = 0.5) +
        geom_violin() + scale_y_log10() +
        theme(legend.position = "none", 
              plot.title = element_text(hjust = 0.5)) +
        ylab("Time (s)") + xlab("Function version") +
        ggtitle("Comparsion of vroom vs fread for 8M entries") +
        scale_x_discrete(labels = c("vroom", "fread")) 
```

Profile an example piece of code:

```{r profile_8M}

profvis({

    #### SNP-level  ##########################################################
    # SNPs annotation and Pathways provided by user
    # all data stored at the WORKSPACE

    ### Load files for analysis
    # all_data <- read_pvals(data_name = "Data/GWASblonde.assoc", 
    #                        snps_ann = "Data/SNP_Table_Annotation_ALL.txt",
    #                        from = "directory")
    
      # Read & format GWAS pvalues
      gwas <- read.table(file = "Data/GWASblonde.assoc",
                         header = TRUE, 
                         stringsAsFactors = FALSE)
      
      ann<- read.table("Data/SNP_Table_Annotation_ALL.txt",
                       header = T, 
                       sep = "\t", 
                       stringsAsFactors = F)
    

    # NOTE: gwas table contains more columns than needed
    # We can select which columns we need (In this case 3,4):
    all_data <- read_pvals(data_name=gwas[,c(3,4)],snps_ann=ann,from="workspace")
    
    # Order data according to the genome
    genome_results <- genome_order(all_data = all_data)

    # Results from genome_order
    ordered_alldata <- genome_results$ordered_alldata
    gs_locs <- genome_results$gs_locs

    # Create new environment to save variables (e.g. pathways, permutations):
    gper.env <- new.env()

    # Pathways can be downloaded using the function get_pathways()
    # Load example pathways into the new environment.
    data(RHSA164843, RHSA446343, RHSA8876384, RHSA8964572,
         RHSA109582, RHSA1474244, envir = gper.env)

    # Map SNPs to pathways
    paths_res <- read2_paths(ordered_alldata = ordered_alldata,
                             gs_locs = gs_locs,
                             sets_from = "workspace",
                             sets_prefix = "RHSA",
                             level = "snp",
                             envir = gper.env)

    # Results from read2_paths:
    pers_ids <- paths_res$per_ors
    pathways <- paths_res$pathways

    # Perform permutations:
    snps_permutation_v2(ordered_alldata = ordered_alldata,
                     pers_ids = pers_ids, 
                     ntraits = 7, 
                     nper = 10, 
                     saveto = "workspace",
                     threshold = 0.05, 
                     gs_locs = gs_locs, 
                     envir = gper.env)

    # Get results
    results <- get_results(res_pattern = "Permus", 
                           level = "snp",
                           from = "workspace", 
                           threshold = 0.05,
                           envir = gper.env)


})

```

Just profile the most expensive routine

```{r test_which}
# is.na returns TRUE for NA values, FALSE for real values
# which returns the position where a TRUE value is found, the test for truth
# is redundant. You can use the is.na directly to return non-na values the
# which may be redundant but will require changes.

# Check
expect_equal(which(is.na(all_data[, 2]) == TRUE),
             which(is.na(all_data[, 2])))

# Test for a performance difference
res <- microbenchmark(
  which(is.na(all_data[, 2]) == TRUE),
  which(is.na(all_data[, 2])),
  is.na(all_data[, 2])
)

# Plot the results
boxplot(res, 
        names = c("which + test", "which", "is.na"),
        col = c("red", "yellow", "green"))

# print the differences
print(res)
```

```{r prof_genome_order}

profvis({
  
    # Order data according to the genome using v2 makes it possible to make
    # changes to the code base to try an improve the performance.
    genome_results <- genome_order_v2(all_data = all_data)

})


```

Check that it still gives the same results.

```{r test_genome_order}

expect_equal(genome_order(all_data = all_data),
             genome_order_v2(all_data = all_data)
             )

```

```{r microbenchmark_genomic_order}

# Run only 5 times as it takes a long time to run
gres <- microbenchmark(
  genome_order(all_data = all_data),
  genome_order_v2(all_data = all_data),
  unit = "s",
  times = 5
)

# print out the results
print(gres)

# Boxplot of the results
boxplot(gres, 
        xlab = "Function", 
        names = c("genome_order", "genome_order_v2"),
        col = c("red", "yellow"))
#
gres %>% select(time, expr)         %>% 
        mutate(time = time / 10^9) %>% # Convert to seconds
        ggplot(aes(x = expr, y = time, fill = expr), alpha = 0.5) +
        geom_violin() + scale_y_log10() +
        theme(legend.position = "none", 
              plot.title = element_text(hjust = 0.5)) +
        ylab("Time (s)") + xlab("Function version") +
        ggtitle("Comparsion of genome_order improvements for 10M entries") +
        scale_x_discrete(labels = c("genome_order", "genome_order_v2")) +
        theme_bw() + theme(legend.position = "none")

```

